### 2020-01-30 paper/educ_work_over_time.tex
Storage for longer discussion of education and conceptual framework

Female Education and Labor Force Participation in India

Female educational attainment has increased substantially over time.
Figure XX shows the distribution of schooling by birth cohort for urban and rural women, 
twenty years or older, whether married or not, based on the four rounds of the NFHS.
The education groups are no education, one to seven years, and eight to eleven years,
and twelve years and above.

@@

Even as the level of female education has increased, the female labor force participation 
in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Afridi2018,Bhargava2018, Chatterjee2018, Bhargava2019}.
A decline in female labor force participation at the beginning of development is
consistent with the hypothesis of a U-shaped labor force participation as a country
develops \citep{Goldin1994}.
India's female labor force participation is, however, lower than most other
countries and more in line with countries in the Middle East and North Africa, and
does not yet show any signs of increasing
\citep{Klasen2015,Chatterjee2018}.

@@

Figure XX shows the percent of married women who are 
currently working at the time of the survey by age group and education level.
No other labor force participation question is consistently available across all 
four surveys. 
Because the question refers to currently working, the percentages will be lower 
than what other studies have found. 
Furthermore, since the question is asked only of married women, the overall female labor 
force participation might have developed differently, especially since many young, 
unmarried women have begun working in, for example, the business process 
outsourcing industry, which, in turn, has increased girls' schooling investments
\citep{Jensen2012}.

@@

Almost all of the urban women who work receive either cash or a combination of 
cash and in-kind for their work (see Appendix Figure XX). 
Rural women have become more likely to receive cash for their labor over time, 
and the better educated, the more likely they are to receive cash.

Despite this, women across all groups have become substantially more likely to 
work for a family member rather than for themselves or somebody else
(see Appendix Figure XX).
Hence, the lower likelihood of working stems mostly from women retracting from 
the general labor market and self-employment.

>>>>

\section{Female Education and Labor Force Participation in India}

Female educational attainment has increased substantially over time.
Figure \ref{fig:education_over_time} shows the distribution of schooling by birth 
cohort for urban and rural women, 
twenty years or older, whether married or not, based on the four rounds of the NFHS.
The education groups are no education, one to seven years, and eight to eleven years,
and twelve years and above.

\begin{figure}[htpb]
\centering
\subfloat[Rural]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_rural}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_urban} 
    \end{minipage}
}
\caption{Distribution of education by cohort for women twenty years or older at survey}
\label{fig:education_over_time}
\end{figure}

@@

Even as the level of female education has increased, the female labor force participation 
in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Afridi2018,Bhargava2018,Chatterjee2018,Bhargava2019}.
A decline in female labor force participation at the beginning of development is
consistent with the hypothesis of a U-shaped labor force participation as a country
develops \citep{Goldin1994}.
India's female labor force participation is, however, lower than most other
countries and more in line with countries in the Middle East and North Africa, and
does not yet show any signs of increasing
\citep{Klasen2015,Chatterjee2018}.

@@

Figure \ref{fig:work_by_survey} shows the percent of married women who are 
currently working at the time of the survey by age group and education level.
No other labor force participation question is consistently available across all 
four surveys. 
Because the question refers to currently working, the percentages will be lower 
than what other studies have found. 
Furthermore, since the question is asked only of married women, the overall female labor 
force participation might have developed differently, especially since many young, 
unmarried women have begun working in, for example, the business process 
outsourcing industry, which, in turn, has increased girls' schooling investments
\citep{Jensen2012}.

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_20} 
    \end{minipage}
}
\caption{Percentage of married women who were working at the time of the 
survey by age group and area of residence}
\label{fig:work_by_survey}
\end{figure}

@@

Almost all of the urban women who work receive either cash or a combination of 
cash and in-kind for their work (see Appendix Figure \ref{fig:work_cash_by_survey}). 
Rural women have become more likely to receive cash for their labor over time, 
and the better educated, the more likely they are to receive cash.

Despite this, women across all groups have become substantially more likely to 
work for a family member rather than for themselves or somebody else
(see Appendix Figure \ref{fig:work_family_by_survey}).
Hence, the lower likelihood of working stems mostly from women retracting from 
the general labor market and self-employment.
